### 2012-10-30 2.tex
Graphs and tables

\noindent
The position that the well-formedness of nonce words is consistent the view of syntactic grammaticality taken in \emph{LSLT} and \emph{Aspects} \citep{LSLT,ASPECTS}, where it is claimed that different syntactic vioaltions result in different degrees of ungrammaticality.

@@

At first blush, this would seem to argue against a naïve account of wellformedness judgements in which ill-formedness results when prosodic parsing fails (e.g., \citealt{Ito1989a}, \citealt{Noske1992}, \citealt{OT}). Consider the possibility that a nonce word is ill-formed if it cannot be syllabified without modification. There is a long history for the proposal that impossible words are those whose surface form cannot be parsed into prosodic structures like syllables without further phonological modification (e.g., \citealt[10f.]{Hooper1973}, \citealt[57f.]{Kahn1976}). There is no doubt that prosodic context is involved in determining where segments may occur. For instance, [ŋ] is a possible coda in English, but not a possible onset. Similarly, the [bn] of \emph{bnick} is judged to be admissible (though unattested) in medial position in English (e.g., \emph{sta}[b.n]\emph{ick}; \citealt[97]{Hooper1973}) but not as an onset. English is unable to sylalbify initial onset /bn/ without further modification, so one might expect underlying /bnɪk/ to be subject to deletion \citep[19f.]{Wolf2009}, prothesis, or anaptyxis. Indeed, \citet{Davidson2006b} finds that English speakers use all three of these repairs when asked to mimic foreign pronunciations of obstruent-nasal onsets impossible in English.\footnote{However, \citet{Davidson2005,Davidson2006a} finds acoustic and articulatory distinctions between the ``exscrescent'' schwa used to resolve non-native onset clusters, and the ``lexical'' schwa found in real words. For example, the anaptyctic [z${^\textrm{ə}}$k] produced by English speakers attempting to mimic word-initial [zk] more closely resembles the onset of \emph{scum} than the start of \emph{succumb}. This may undermine any link between non-native realizations and epenthesis of schwa in native words (e.g., in the syllabic allomorphs of the regular past tense and noun plural).}

However, prosodic well-formedness distinctions do not provide a ready explanation for the ``numerous other distinctions'' posited in \emph{SPE}. This is thought to rule out simple prosodic account of wordlikeness judgements.

\begin{quote}
A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements. \citep[371]{Shademan2006}
\end{quote}

Implicit in this claim is the assumption that the graded judgements obtained from wordlikeness tasks which permit gradient responses require a gradient model of well-formedness. However, it is argued here this is not pertinent to the question of whether wordlikeness is gradient or categorical, because graded judgements are a property of all graded tasks, even when applied to judgements of discrete categories.

\citet[382]{Hayes2008a} ``consider the ability to model gradient intuitions to be an important criterion for evaluation phonotactic models''. However, they fail to ask consider whether categorical models provide a reasonable fit to gradient judgements. Rather than directly evaluating the categorical model of English phonotactics proposed by \citet{Clements1983}, \citeauthor{Hayes2008a} convert these constraints, many of which are exceptionless, into probabilitistic constraints. This suggests that \citeauthor{Hayes2008a} equate ``the ability to model gradient intuitions'' with the ability to make continuous predictions about well-formedness. But if gradient judgements are nothing more than a task effect, a model need not make gradient predictions to fit gradient judgements.

Despite many claims that categorical models are empirically insufficient, there is not a single quantitative investigation of categorical wordlikeness models, and no study which compares categorical and gradient wordlikeness models on an equal footing. This chapter represents a first attempt to fill that lacuna.

OUTLINE HERE

\section{What wordlikeness might not be}

\subsection{Gradience in wordlikeness judgements}

\subsection{Gradience as a task effect}

\section{Evaluation} 
\label{2evaluation}

\subsection{Method}

\subsubsection{Materials and subjects}

\subsubsection{Computational models}

\subsection{Results}

\subsection{Discussion}

\section{Conclusions}

%\citet{Armstrong1983} cut the Gordian knot by asserting the possibility of observing a false consequent; they find the existence of all-or-nothing concepts like ``odd number'' or (more controversially) ``female'' to be apparent.
%\begin{quote}
%Are there definitional concepts? Of course. For example, consider the concept \emph{odd number}. This seems to have a clear definition, a precise description. [\ldots{}] No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
%\end{quote}
%This proposition has the form of an indicative conditional. While the antecedent of this condition, the presence or absence of intermediate ratings in a graded judgement task, is readily observable, it has no bearing on the truth or falsity of the conditional proposition. However, whereas a false consequent (e.g., an all-or-nothing concept) would potentially falsify the proposition, the consequent, a mental state, cannot be directly observed. If this is the case, then both the proposition and the consequent (the proposition that wordlikeness is itself gradient) fail the test of falsification, placing it beyond the scope of scientific inquiry.
%\subsection{Gradience as a liability}
%One might reasonably ask whether 
%representations 
%\citeauthor{Armstrong1983} note that if subjects produce intermediate ratings for these categories, then the naïve proposition is false, and this is what \citeauthor{Armstrong1983} find in their experiments. They ask subjects to rate, on a seven-point scale, the extent to which, e.g., certain odd counting numbers represent the concept ``odd number'', and find that speakers do make use of intermediate values, as shown in Figure \ref{agg}. In summary, the proposition that intermediate ratings indicate mental gradience is either untestable, or it is false, and is thus rejected.
%\begin{figure} 
%\centering
%\includegraphics{agg.pdf}
%\caption{Subjects asked to rate the degree to which certain even and odd numbers represent the categories ``even'' and ``odd'', respectively, freely use intermediate ratings.}
%\label{agg} 
%\end{figure}
%\citeauthor{Armstrong1983} do not view their results as evidence that subjects have an internal model of odd numbers which is at odds with the formal, all-or-nothing definition, and in fact, several additional experiments show that subjects deploy the extension of the formal definition in other tasks. Reviewing this study, \citet[215]{Schutze2011} writes that the experiments show that judgements are ``sensitive to factors other than our underlying competence''. In the case of wordlikeness, a case can be made that one such factor is similarity to existing words (e.g., \citealt{Bailey2001}, \citealt[][151, fn. 27]{LSLT}, \citealt{Greenberg1964}, \citealt{Ohala1986b}, \citealt{Schutze2005}, \citealt{Sendlmeier1987}, \citealt{Vitz1973}). However, the mere fact that subjects use intermediate ratings does not show that they do so with any systematicity, or that, e.g., the contrast between \emph{447} and \emph{7} with respect could be given any satisfying explanation. \citeauthor{Schutze2011} further suggests that gradient responses are the product of the task itself:
%\begin{quote}
%Putting it another way, when asked for gradient responses, participants will find some way to oblige the experimenter; if doing so is incompatible with the experimenter's actual question, they apparently infer that she must have really intended to ask something slightly different. \citep[215]{Schutze2011}
%\end{quote}
%This introduces a troubling possibility, that subjects ``oblige'' experimenters by introducing random noise to their categorical judgements when presented with a Likert task. Were this the case, the observed intermediate judgements are unlikely to yield information about speakers' knowledge of possible and impossible words. The alternative considered here is that there exists some model in which the gradience in ratings is systematic. 
%The evaluation makes exclusive use of previously published English wordlikeness data. For inclusion in the evaluation, a study must conform to all three of the following conditions. First, the stimuli must consist only of monosyllabic words presented auditorily. Secondly, a significant portion of the stimuli must contain gross phonotactic violations (e.g., \emph{bnick}). Finally, the ratings, averaged across subjects, must be publicly available. The large-scale studies by \citet{Bailey2001} and \citet{Shademan2006,Shademan2007} are ineligible because the former lack stimuli with gross phonotactic violations and the latter data are not available to the public in any form. All data are reproduced in Appendix \ref{appendixA}.
%\subsubsection{\citealt{Scholes1966}}
%\citet{Scholes1966} conducted a number of English wordlikeness judgement with middle school children. The data used here come from his ``experiment 5'', in which 63 monosyllabic items were presented to 33 seventh-grade students. For each stimulus, the subjects produced forced choice ``yes''/``no'' answers to the question of whether the item ``is likely to be usable as a word of English''. \citet{Hayes2008a} and \citet{Albright2009a} analyze this data as gradient by performing an item averaging using the fraction of ``yes'' answers for each stimulus \citep[see also][]{Pierrehumbert1994,Coleman1997,Frisch2000}. For instance, 22 of the 33 students answered ``yes'' for \emph{shlerk} [ʃlɚk], so it is assigned a score of $0.666$.\footnote{This procedure conflates intraspeaker variation, which in some cases is considerable \citep{Shademan2007} with speaker-internal gradience, and its adoption here should not be construed as an endorsement. Rather, it is provided solely for comparison with prior studies using the \citeauthor{Scholes1966} data.} These stimuli are all ostensibly non-words, but include \emph{clung} [klʌŋ], the preterite and past participle of the verb \emph{cling}, and \emph{brung} [brʌŋ], a dialectical past participle for \emph{bring}. These two words were excluded and the remaining 61 stimuli were submitted to analysis. 
%\subsubsection{\citealt{Albright2003b}}
%\citet{Albright2003b} gathered wordlikeness judgements to serve as norms for a \emph{wug}-test. 87 items were presented to 24 undergraduate subjects, who rated each word on a seven-point Likert scale. The lowest point on the scale was labeled ``completely bizarre, impossible as an English word'', and that the highest point was labeled ``completely normal, would make a fine English word''.
%\begin{table}
%\centering
%\begin{tabular}{l r r r}
%\toprule
%                        & subjects & items & trials \\
%\midrule
%\citealt{Albright2003a} & 24       & 87    & 2,088  \\
%\citealt{Albright2007}  & 68       & 40    & 2,720  \\
%\citealt{HayesInPress}  & 29       & 80    & 2,320  \\
%\citealt{Scholes1966}   & 33       & 63    & 2,178  \\
%\midrule
%\textsc{Total}          & 154      & 270   & 9,306  \\
%\bottomrule
%\end{tabular}
%\caption{Subject and item counts}
%\label{counts}
%\end{table}
%\subsection{Model comparison}
%The four models used here consist of a binary baseline and three computationally implemented gradient models. These three models are chosen because prior studies have shown they are correlated with wordlikeness judgements. Non-parametric rank correlation statistics are used evaluate the correlation between model predictions and wordlikeness ratings. Rank correlation are used rather than the parametric Pearson correlation statistic that is used by \citet{Hayes2008a}, for instance, because they make none of the potentially troublesome assumptions of the latter method \citep[see][23, fn. 12]{Albright2009a}.
%The Spearman $\rho$ is most widely known rank correlation statistic, but it is difficult to give a natural interpretation to this quantity. On the other hand, the Kendall $\tau_b$ and Goodman-Kruskal $\gamma$ can be interpreted as fractions of the number of \emph{concordant} and \emph{discordant} pairs \citep{Noether1981}. Consider the case here, in which model scores are compared with wordlikeness ratings. If a model rates \emph{dresp} [dɹɛsp] more wordlike than *\emph{srest} [sɹɛst] and if speakers further rate \emph{dresp} more English-like than \emph{srest}, than \emph{dresp}/\emph{srest} are a concordant pair. If, however, a model and speakers disagree on the relative ranking of \emph{dresp} and \emph{srest}, the pair is discordant. The $\tau_b$ and $\gamma$ differ only in their treatment of ``ties'' (i.e., if e.g., *\emph{dresp} and *\emph{srest} are scored the same, or rated the same); $\tau_b$ includes a correction for ties, whereas $\gamma$ ignores tied pairs. Much like the familiar Pearson correlation, $\rho$, $\tau_b$ and $\gamma$ are all in the range [$-1$, $1$]. Correlations of the four models are given in Table \ref{cor}.
%\subsubsection{Binary baseline}
%The binary baseline used here is a crude implementation of the null hypothesis that there are no gradient effects in wordlikeness judgements. To create such a baseline, it is necessary to distinguish between nonce words which contain a gross phonotactic violation and those which do not. As all stimuli here are monosyllables, this task can be further simplified by separately considering the two major subcomponents of the syllable, the onset and rime. Speakers are particularly adept as separating onsets from rimes \citep{Treiman1986,Treiman1995,Fowler1993}, and a large portion of phonotactic generalizations and violations can be localized to either onset or rime \citep[e.g.,][]{Borowsky1989,Fudge1969,Treiman1995,Kessler1997,Treiman2000}.
%The baseline considers a nonce syllable to be well-formed if it consists of both an attested onset and an attested rime; the free combination of these two components is the only mechanism by which this model can generalize beyond attested words. It is surely the case that this is an ultimately insufficient model: \citet{Albright2009a} observes, for instance, that [ɛsp] is a well-formed rime, even though it is found in no word of English. This problem is more acute in languages with more permissive syllable structures than those of English. For instance, \citet{Fischer-Jorgensen1952} and \citet{Vogt1954} assert that there are many accidental gaps (i.e., possible but unattested structures) in the inventory of consonant clusters in Georgian, a language which admits as many as five adjacent consonants.\footnote{
%Chapter \ref{clusters} 
%%\citet{Gorman2012c}
%reports a similar result for English syllable contact clusters.} Despite these defects, the baseline outperforms the gradient models in most contexts.
%The rating densities from the three studies, linearly transformed to fit within the interval [0, 1] and split according to this binary baseline, are shown in Figure \ref{dsn}. In all three studies, it is possible to discern a relatively sharp separation between valid and invalid clusters, but also the presence of intermediate values.
%\begin{figure} 
%\centering
%\includegraphics{density.pdf}
%\caption{Average ratings of individual nonce words, linearly transformed to fit the interval [0, 1], tend to clump into two groups with little overlap; words which consist of  attested onsets and rimes receive ratings near ceiling, whereas ratings of phonotactically invalid words are spread across the lower half of the spectrum.}
%\label{dsn}
%\end{figure}
%\begin{table} 
%\centering
%\begin{tabular}{l l l l l}
%\toprule
%Spearman $\rho$          & baseline         & maxent  & bigram           & density          \\
%\midrule
%\citealt{Greenberg1964}  & {0.845}          & {0.765} & {\textbf{0.863}} & {0.648}          \\
%\citealt{Scholes1966}    & {0.791}          & {0.762} & {\textbf{0.827}} & {\textbf{0.827}} \\
%\citealt{Albright2003b}  & {0.725}          & {0.429} & {0.708}          & {\textbf{0.742}} \\
%\midrule
%Kendall $\tau_b$         & baseline         & maxent  & bigram           & density          \\
%\midrule
%\citealt{Greenberg1964}  & {\textbf{0.716}} & {0.585} & {0.692}          & {0.462}          \\
%\citealt{Scholes1966}    & {\textbf{0.664}} & {0.597} & {0.652}          & {0.565}          \\
%\citealt{Albright2003b}  & {\textbf{0.599}} & {0.343} & {0.506}          & {0.556}          \\
%\midrule
%Goodman-Kruskal $\gamma$ & baseline         & maxent  & bigram           & density          \\
%\midrule
%\citealt{Greenberg1964}  & {\textbf{1.000}} & {0.684} & {0.692}          & {0.462}          \\
%\citealt{Scholes1966}    & {\textbf{0.995}} & {0.634} & {0.667}          & {0.614}          \\
%\citealt{Albright2003b}  & {\textbf{0.953}} & {0.656} & {0.509}          & {0.575}          \\
%\bottomrule
%\end{tabular}
%\caption{Rank correlations between wordlikeness ratings and phonotactic models surprisingly reveal that the binary baseline meeds or exceeds the coverage of three state-of-the-art phonotactic models. All correlations are significant at $p = 0.05$.}
%\label{cor}
%\end{table}

@@

%\begin{table} 
%\centering
%\begin{tabular}{l r r r r}
%\toprule
%                         & \multicolumn{1}{c}{featural bigrams} & \multicolumn{2}{c}{segmental bigrams}  \\
%                         & maximum likelihood                   & maximum likelihood & Laplace smoothing \\
%\midrule
%%Pearson $\rho
%Spearman $\rho$          & 0.638                                & 0.660              & \textbf{0.708}    \\
%Kendall $\tau_b$         & 0.457                                & 0.467              & \textbf{0.506}    \\
%Goodman-Kruskal $\gamma$ & 0.462                                & 0.473              & \textbf{0.509}    \\
%\bottomrule
%\end{tabular}
%\caption{Segmental bigrams outperform ``featural bigrams'', and Laplace smoothing increases the correlation between the segmental bigram model proposed by \citet{Albright2009a}, which uses maximum likelihood estimation, and the \citet{Albright2003b} wordlikeness norms. All correlations are significant at $p = 0.05$.}
%\label{albrightimproved}
%\end{table}

@@

%\begin{table} 
%\centering
%\begin{tabular}{l r r r}
%\toprule
%Kendall $\tau_b$          & maxent         & bigram         & density  \\
%\midrule
%\citealt{Greenberg1964}   & 0.670          & \textbf{0.680} & 0.501 \\
%\citealt{Scholes1966}     & \textbf{0.685} & 0.632          & 0.639 \\
%\citealt{Albright2003b}   & 0.542          & 0.603          & \textbf{0.623} \\
%\bottomrule
%\end{tabular}
%\caption{The binary baseline is strongly correlated with the three gradient model scores; all correlations are significant at $p = 0.05$.}
%\label{bcor}
%\end{table}
%\begin{table} 
%\centering
%\begin{tabular}{l r r r}
%\toprule
%$\Delta$ Spearman $\rho$          & maxent            & bigram            & density  \\
%\midrule
%\citealt{Greenberg1964}  & $-0.060$          &  \textbf{0.038} & $-0.017$ \\
%\citealt{Scholes1966}    & $-0.029$          &  \textbf{0.047} & $-0.035$ \\
%\citealt{Albright2003b}  & $-0.008$          & $-0.015$          & \textbf{0.018} \\
%\midrule
%$\Delta$ Kendall $\tau_b$         & maxent            & bigram            & density  \\
%\midrule
%\citealt{Greenberg1964}  & $-0.114$          & \textbf{$-$0.007} & $-0.084$ \\
%\citealt{Scholes1966}    & $-0.067$          & \textbf{0.003}  & $-0.061$ \\
%\citealt{Albright2003b}  & \textbf{$-$0.038} & $-0.092$          & $-0.049$ \\
%\midrule
%$\Delta$ Goodman-Kruskal $\gamma$ & maxent            & bigram            & density  \\
%\midrule
%\citealt{Greenberg1964}  & $-0.268$          & \textbf{$-$0.260} & $-0.337$ \\
%\citealt{Scholes1966}    & $-0.361$          & \textbf{$-$0.313} & $-0.345$ \\
%\citealt{Albright2003b}  & \textbf{$-$0.137} & $-0.443$          & $-0.386$ \\
%\bottomrule
%\end{tabular}
%\caption{The change in rank correlation generated by augmenting the purely binary model with gradient predictions is small and in most cases it is negative.}
%\label{controlled}
%\end{table}

>>>>

\noindent
The position that the well-formedness of nonce words is consistent the view of syntactic grammaticality taken in \emph{LSLT} and \emph{Aspects} \citep{LSLT,ASPECTS}, where it is claimed that different syntactic violations result in different degrees of ungrammaticality.

@@

\begin{quote}
\ldots{}A defect of current grammatical acounts of phonotactics is that they render simple up-or-down decisions concerning well-formedness and cannot account for gradient judgements. \citep[371]{Shademan2006}
\end{quote}

\citet[382]{Hayes2008a} ``consider the ability to model gradient intuitions to be an important criterion for evaluating phonotactic models''.

\begin{quote}
No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
\end{quote}

\begin{quote}
Some have responded to these findings very consistently, by asserting that the experimental findings are to be interpreted as before: that, psychologically speaking, odd numbers as well as birds and vegetables are graded concepts\ldots{} We reject this conclusion just because we could not explain how a person could compute with integers who believed that 7 was odder than 23. We assert confidently that the facts about subjects being able to compute and about their being able to give the definition of odd number, etc., are the more important, highly entrenched, facts we want to preserve and explain\ldots{} we ourselves are prepared to give up the seeming fact that some odd numbers appear, as shown by their behavior in certain experimental paradigms, to be odder than others\ldots{}we do not give it up by saying that it was no fact; rather, by saying it must have been a fact about something other than the structure of concepts. \citep[284]{Armstrong1983}
\end{quote}

\begin{quote}
\ldots{}we hold that \emph{fruit} and \emph{odd number} have different structures, and yet we obtain the same experimental outcome for both. But if the same result is achieved regardless of the concept struture, then the experimental design is not pertinent to the determination of concept structure. \citep[284--5]{Armstrong1983}
\end{quote}

At first blush, this would seem to argue against a naïve account of wellformedness judgements in which ill-formedness results when prosodic parsing fails (e.g., \citealt{Ito1989a}, \citealt{Noske1992}, \citealt{OT}). Consider the possibility that a nonce word is ill-formed if it cannot be syllabified without modification. There is a long history for the proposal that impossible words are those whose surface form cannot be parsed into prosodic structures like syllables without further phonological modification (e.g.,

\begin{figure}
\centering
\includegraphics{density.pdf}
\caption{Average ratings of individual nonce words, linearly transformed to fit the interval [0, 1]}
\label{dsn}
\end{figure}

\begin{table}
\centering
\begin{tabular}{l rrr}
\toprule
                        & subjects & items & trials \\
\midrule
\citealt{Albright2007}  & 68       & 40    & 2,720  \\
\citealt{Albright2003a} & 24       & 87    & 2,064  \\
\citealt{Scholes1966}   & 33       & 63    & 2,178  \\
\midrule
\textsc{Total}          & 125      & 187   & 6,962  \\
\bottomrule
\end{tabular}
\caption{Subject and item counts}
\label{counts}
\end{table}

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
Pearson $r$               & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
\midrule
\citealt{Albright2007}    & \textbf{.725}    & {.703}          & {.463}          & {.695}           \\
\citealt{Albright2003b}   & {.594}           & {.208}          & \textbf{.746}   & {.688}           \\
\citealt{Scholes1966}     & {.803}           & {.534}          & {.737}          & \textbf{.834}    \\
\midrule
Spearman $\rho$           & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
\midrule
\citealt{Albright2007}    & \textbf{.819}    & {.661}          & {.338}          & {.612}           \\
\citealt{Albright2003b}   & {.662}           & {.389}          & {.699}          & \textbf{.739}    \\
\citealt{Scholes1966}     & {.769}           & {.575}          & {.794}          & \textbf{.817}    \\
\midrule
Kendall $\tau_{b}$        & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
\midrule
\citealt{Albright2007}    & {.666}           & \textbf{.683}   & {.245}          & {.450}           \\
\citealt{Albright2003b}   & {.474}           & {.159}          & {.499}          & \textbf{.558}    \\
\citealt{Scholes1966}     & {.619}           & {.484}          & {.614}          & \textbf{.671}    \\
\midrule
Goodman-Kruskal $\gamma$  & \textsc{prosody} & \textsc{maxent} & \textsc{bigram} & \textsc{density} \\
\midrule
\citealt{Albright2007}    & \textbf{.874}    & {.849}          & {.502}          & {.574}           \\
\citealt{Albright2003b}   & \textbf{.929}    & {.606}          & {.246}          & {.488}           \\
\citealt{Scholes1966}     & \textbf{.907}    & {.558}          & {.628}          & {.738}           \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{scores}
\end{table}

@@

@@

### 2012-10-30 3.tex
Graphs and tables

\chapter{Static and derived phonotactic preferences in Turkish}
\label{turkish}

@@

Two less familiar notations are used in this chapter. First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules. Directional application provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}). Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.

@@

While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show that Turkish speakers internalize the tendency for roots to conform to backness harmony. If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it. Several other ``external'' facts suggest that this is not the case. The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.

The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} Some speakers are said to be capable of pronouncing these non-native clusters, but in fast speech the cluster is split by anaptyxis. In most cases, this vowel matches the following root vowel for backness.

@@

Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} 
The game is native to the related language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; it is not indigenous to Turkish, but can be taught quickly to children or adults. In this game, the base is reduplicated and the first vowel of the reduplicant replaced with a [$+$\textsc{Back}] vowel. In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the inserted [$+$\textsc{Back}] vowel. The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}

@@

\noindent 
In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant. On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{There is reason to believe that reduplicants bear lexical diacritics. Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, provides such a case. Verb reduplication requires a bisyllabic reduplicant. This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots actually have three copies of the root, and since reduplicated forms of many trisyllable verbs are ineffable. However, a few trisyllabic verbs exceptionally show full reduplication. Exceptionality appears to be a lexical property (since it is specific to certain lexical entries), but it surfaces only in the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.

@@

\section{Evaluation}

@@

\citet{Zimmer1969} administers two variants of the paired nonce word rating task. The first used 23 native adult speakers who were permitted to select either nonce word as more like Turkish, or to indicate `no preference'. For the purposes below, `no preference' results are discarded. The second experiment used 32 native adults, none of whom appeared in the preceding study, and used a forced binary choice with some minor changes to the stimulus set.

>>>>

\chapter{Static and derived Turkish phonotactics}
\label{turkish}

@@

Two less familiar notations are used in this chapter. First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules. Since it was first proposed by \citet{Johnson1972}, considerable evidence for directional application has been adduced (e.g., \citealt{Anderson1974}: chap.~9, \citealt{Gorman2012e}, \citealt{Howard1972}:65f., \citealt{Kavitskaya2008}, \citealt{Kaye1982}, \citealt{Kenstowicz1977}:189f., \citeyear{Kenstowicz1979}:319f., \citealt{Piggott1975}, \citealt{Sohn1971}; see \citealt{McCarthy2003b} and \citealt{Wolf2011b} for recent reviews.) Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.

@@

While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show that Turkish speakers internalize the tendency for roots to conform to backness harmony. If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it. Several other ``external'' facts suggest that this is not the case. The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.\footnote{Thanks to Bert Vaux and Kie Zuraw for bringing these studies to my attention.}

The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}. Some speakers are said to be capable of pronouncing these non-native clusters, but in fast speech the cluster is split by anaptyxis. In most cases, this vowel matches the following root vowel for backness.

@@

Similar evidence comes from a language game discussed by \citet{Harrison2001}. The game is native to the related language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; it is not indigenous to Turkish, but can be taught quickly to children or adults. In this game, the base is reduplicated and the first vowel of the reduplicant replaced with a [$+$\textsc{Back}] vowel. In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the inserted [$+$\textsc{Back}] vowel. The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}

@@

\noindent 
In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant. On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{There is reason to believe that reduplicants bear lexical diacritics. Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, provides such a case. Verb reduplication requires a bisyllabic reduplicant. This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots actually have three copies of the root, and since reduplicated forms of many trisyllable verbs are ineffable. However, a few trisyllabic verbs exceptionally show full reduplication. This is a lexical property, but one that is only seen on the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.

@@

\section{Evaluation}
\label{3evaluation}

@@

\citet{Zimmer1969} administers two variants of the paired nonce word rating task. The first used 23 native adult speakers who were permitted to select either nonce word as more like Turkish, or to indicate `no preference'. For the purposes below, `no preference' results are discarded. The second experiment used 32 native adults, none of whom appeared in the preceding study, and used a forced binary choice.
